### 2014-04-24 diagrams/bondPair.tex

\begin{tikzpicture}[y=0.80pt, x=0.8pt,yscale=-1, inner sep=0pt, outer sep=0pt]
\begin{scope}[shift={(-71.10332,-65.91978)}]
  \path[draw=black,dash pattern=on 1.60pt off 0.80pt,line join=miter,line
    cap=butt,miter limit=4.00,line width=0.800pt] (100.0000,102.3622) --
    (200.0000,102.3622) -- (300.0000,102.3622);
  \path[draw=black,line join=miter,line cap=butt,line width=0.800pt]
    (72.0000,92.3622) -- (75.0000,77.3622) -- (200.0000,102.3622) --
    (300.0000,77.3622) -- (305.0000,97.3622);
  \path[draw=black,line join=miter,line cap=butt,line width=0.800pt]
    (75.0000,77.3622) -- (200.0000,102.3622) -- (300.0000,77.3622);
  \path[draw=black,line join=miter,line cap=butt,line width=0.800pt]
    (200.0000,102.3622) -- (198.0000,67.3622);
  \path[fill=black] (178.797,148.02312) node[above right] (text6246) {};
  \path[fill=black] (276.10333,95.203163) node[above right] (text6273)
    {$\mathbf{\xi}$};
  \path[fill=black] (106.10332,125.20316) node[above right] (text6277)
    {$\mathbf{\xi}$};
  \path[fill=black] (146.10332,75.203163) node[above right] (text6281)
    {$\vstate{Y}{}{\mathbf{-\xi}}$};
  \path[fill=black] (256.10333,80.203163) node[above right] (text6285)
    {$\vstate{Y}{}{\mathbf{\xi}}$};
\end{scope}

\end{tikzpicture}